## Effects of Land Use on Soil Moisture

```{r}
#Load packages
require(ggplot2)
inst
library(ggplot2)
library(GGally)
library(lme4)
library(dplyr)
library(lmerTest)
```


```{r}
soil <- read.csv("soil.csv")
soil <- soil %>%
  #select(-row, -stake) %>%
  rename %>%
  mutate(soil = case_when(
    soil == "FF" ~ "Forest floor",
    soil == "MIN" ~ "Mineral soil",
    TRUE ~ soil)) %>%
  mutate(site = case_when(
    site == "P1" ~ "Plow 1",
    site == "P2" ~ "Plow 2",
    site == "S1" ~ "Pasture 1",
    site == "S2" ~ "Pasture 2",
    site == "W1" ~ "Woodlot 1",
    site == "W2" ~ "Woodlot 2",
    TRUE ~ site))

#Remove NAs
soil <- na.omit(soil)
#soil <- drop_na(soil$soil.moisture)
soil$stake <- as.factor(soil$stake)
soil$row <- as.factor(soil$row)
soil$site <- as.factor(soil$site)

#Visualize data
ggpairs(soil, c("site", "row", "stake", "soil", "soil.mass", "soil.moisture", "soil.moisture.capacity"))

#Count unique values for each variable
sapply(lapply(soil, unique), length)
  #soil variable only has one unique value (Forest floor) --> drop from model

#Create a FULL model --> soil moisture as dependent variable
smois1 <- lmer(soil.moisture ~ factor(site) + (1|row) + (1|stake) + ff.thickness + bulk.density + soil.mass + ph.h2o + ph.cacl2 + c + n + om + p + nh4 + no3 + n.min + nitr + cec + ca + mg + k + soil.moisture + soil.moisture.capacity, data = soil)
summary(smois1)
plot(smois1)
  #Random effects
    #standard deviation -->  This is a measure of the variability for each random effect that you added to the model
    #row, stake -->effect of row/stake, overall variation within row/stake, gives us the variance...std. dev. is the effect size (effect of row/stake on overall soil moisture)
   #Residual --> stands for the variability that’s not due to either stake or row
    #row effect is greater than stake --> looking at std. dev. to determine this
  #Fixed effects
    #coefficients --> represent the slope for the categorical effect of soil data
    #t- value --> simply the estimate divided by the standard error
    #intercept = the average of our data for the Pasture 1 location condition

smois2 <- update(smois, .~. -bulk.density)
summary(smois2)

smois3 <- update(smois2, .~. -cec)
summary(smois3)

smois4 <- update(smois3, .~. -n.min)
summary(smois4)

smois5 <- update(smois4, .~. -nitr)
summary(smois5)

smois6 <- update(smois5, .~. -k)
summary(smois6)

smois7 <- update(smois6, .~. -c)
summary(smois7)

smois8 <- update(smois7, .~. -no3)
summary(smois8)

smois9 <- update(smois8, .~. -p)
summary(smois9)

smois10 <- update(smois9, .~. -om)
summary(smois10)

smois11 <- update(smois10, .~. -ph.cacl2)
summary(smois11)

smois12 <- update(smois11, .~. -ca)
summary(smois12)

smois13 <- update(smois12, .~. -mg)
summary(smois13)

smois14 <- update(smois13, .~. -nh4)
summary(smois14)

smois15 <- update(smois14, .~. -ph.h2o)
summary(smois15)

smois16 <- update(smois15, .~. -soil.mass)
summary(smois16)

smois17 <- update(smois16, .~. -n)
summary(smois17)

#Compare models with AIC
AIC(smois1, smois2, smois3, smois4, smois5, smois6, smois7, smois8, smois9, smois10, smois11, smois12, smois13, smois14, smois15, smois16, smois17)
  #smois17 is the best model

#Assumptions: creating a residual plot, a histogram of the residuals or a Q-Q plot 

#Best Model
smois <- lmer(soil.moisture ~ factor(site) + (1|row) + (1|stake) + ff.thickness + soil.moisture.capacity, data = soil)
summary(smois)

par(mfrow = c(2,2))
plot(smois)

library(car)
vif(smois)

library(ggplot2)
#Site vs soil moisture (scatterplot)
ggplot(soil, aes(x=factor(site), y=soil.moisture)) +
  geom_point() +
  geom_smooth(se = FALSE, method = lm) +
  theme_bw() 

library(ggplot2)
#Site vs soil moisture (boxplot)
ggplot(soil, aes(x=factor(site), y=soil.moisture)) +
  geom_boxplot() +
  geom_smooth(se = FALSE, method = lm) +
  theme_bw() + 
  labs(x="Site (Land Use Type)", y="Soil Moisture (gH20/gsoil)")

#Mean of soil moisture for each of the land use types
soil %>% filter(site == "Plow 1" | site == "Plow 2") %>%
  summarise(mean=mean(soil.moisture))
  #0.9737

soil %>% filter(site == "Pasture 1" | site == "Pasture 2") %>%
  summarise(mean=mean(soil.moisture))
  #1.238039	

soil %>% filter(site == "Woodlot 1" | site == "Woodlot 2") %>%
  summarise(mean=mean(soil.moisture))
  #1.604906	
```

## Writeup

Null hypothesis: There is no significant relationship between any of the soil variables on soil moisture (gH20/gsoil). 
Alternative hypothesis: This is a significant relationship between one/several of the soil variables on soil moisture (gH20/gsoil). 

We first wanted to understand if soil moisture (gH20/gsoil) is influenced by any of the soil variables provided in this data set. To do so, we created a linear model ('lmer') with soil moisture as the response variable and all the remaining variables as the independent variables. Row and stake were treated as random effects, and site was treated as a factor. The full model was reduced down to just contain significant variables, which included: site, row (random effect), stake (random effect), forest floor thickness (), and soil moisture capacity. Based on the results of the model 


